---
title: "processing"
author: "Jamie Cummins"
date: "3/15/2021"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(Hmisc)
library(RSQLite)
library(jsonlite)
library(tidyverse)
library(janitor)
library(pacman)
library(knitr)
library(kableExtra)
library(lubridate) 
library(RCurl)
library(threadr)
library(emayili)

```


# download data from server and save to specific location 

```{r}

## Change these:
### make sure you leave the "data.sqlite" and "data.csv" bits
download_location <- paste0('raw/training/smart_data_', format(Sys.Date(), "%d%m%Y"), '.sqlite')
save_location <- paste0('processed/smart_data_', format(Sys.Date(), "%d%m%Y"), '.csv')


## Do not change these:
url <- "ftp://ftp.drjamiecumminscom.webhosting.be/www/md/training-1/data/data.sqlite"
credentials <- "mariedelabie@drjamiecumminscom:MarieDelabie1"



# download questionnaire data and save it to location
download_ftp_file(file_remote = url,
                  file_local = download_location,
                  credentials = credentials)

```

# Converting SQL data

Establish SQL connection and extract main table:

```{r}

# Establish SQL connection
con <-
  dbConnect(drv = RSQLite::SQLite(),
            # GUYS: CHANGE THE BELOW DIRECTORY TO THE LOCATION AND NAME OF THE FILE
            dbname = download_location)


# Extract main table
d <-
  dbGetQuery(conn = con,
             statement = 'SELECT * FROM labjs')

# Close and discard connection
dbDisconnect(conn = con)
rm(con)

```

* id: row number
* session: generated identifier for each individual session (one per participant)
* timestamp
* url
* metadata: also includes a variable "id" which corresponds to individual sessions
* data: the actual data

Process metadata and combine with the rest:

```{r}

d.meta <-
  map_dfr(d$metadata, fromJSON) %>% # apply JSON function to the metadata
  dplyr::rename(unique_id = id) # rename to avoid confusion

d <- d %>%
  bind_cols(d.meta) %>%
  select(-metadata)

```

* slice: the number of the entry for that specific subject
* observation: what we will use as subject identifier
* payload: tells you whether the data are complete (full) or not (incremental)

See by how much subject identifiers can be shortened:

```{r}

count_unique <- function(x) {
  return(length(unique(x)))
}
# returns how many unique observations there are in a given vector

information_preserved <- function(x, length) {
  return(count_unique(str_sub(x, end = i)) == count_unique(x))
}
# put in a vector and a string length
# for that string length, the vector checks how many unique observations there are when the strings are truncated using that length
# the function then compares this number to the number of unique observations when at full string length
# if this number is the same (i.e., truncating the string doesn't change the number of unique observations) it returns TRUE

for (i in 5:36) {
  if (
    information_preserved(d$session, i) &&
    information_preserved(d$observation, i)
  ) {
    break()
  }
}
# check if more than five characters are needed to preserve the identifying information
# for-loop breaks as soon as both expressions are evaluated as TRUE

```

Truncate ID strings to five characters:

```{r}

# Truncate IDs
d <- d %>%
  dplyr::mutate(session = str_sub(session, end = i),
                unique_id = str_sub(unique_id, end = i))

```

Extract the actual data:

```{r}

# Create function that does this
parseJSON <- function(input) {
  return(input %>%
           fromJSON(flatten = TRUE) %>% {
             if (class(.) == 'list') {discard(., is.null) %>% as_tibble()}
             else {.}
             } %>%
           janitor::clean_names() %>%
           mutate_all(as.character))
}

# Get full data
d.full <- d %>%
  dplyr::filter(payload == 'full') %>%
  group_by(unique_id, id) %>%
  do({map_dfr(.$data, parseJSON)} %>%
       bind_rows()) %>%
  ungroup()

```

Get incremental data (in case the full data are not available).

```{r}

d.incremental <-
  d %>%
  dplyr::filter(payload %in% c('incremental', 'latest')) %>%
  group_by(unique_id, id) %>%
  do(
    {map_dfr(.$data, parseJSON)} %>%
      bind_rows()
  ) %>%
  ungroup()

# # merge data sets
# d.output <-
#   d.full %>%
#   bind_rows(d.incremental %>% filter(!(unique_id %in% d.full$unique_id)))

# %>%
#   type_convert()

full_raw_data_df <- d.incremental

```

# write out and save csvs

```{r}

# CHANGE LOCATION AND NAME IN THE STRING TO BE WHERE YOU WANT TO SAVE THE FILE
write_csv(full_raw_data_df, save_location)

```


Notes to self:

1. Go through SMART data -> identify participants who completed only partial training but also gave follow-up measures, then run sensitivity analyses
2. Extract descriptive data about the number of sessions that each participant completed
3. flag and !!probably!! include participants who only completed T1 assessments -> sensitivity analysis
4. Check for bugs in the exclusion of participants at T2
5. Check out the potential for random responding in RAI/other measures


# process demographics

```{r}

# questionnaire_df <- questionnaire_raw_df %>%
#   filter(!is.na(wat_is_uw_e_mailadres)) %>%
#   select(email = wat_is_uw_e_mailadres,
#          age = wat_is_uw_leeftijd,
#          complaints = welke_cognitieve_klachten_heeft_u,
#          preexisting_condition = hebt_u_een_bepaalde_neurologische_of_psychiatrische_diagnose,
#          brain_injury = hebt_u_ooit_een_hersenletsel_gehad)

```


```{r}

# write_csv(questionnaire_df, paste0("processed/questionnaire_processed_data_", format(Sys.Date(), "%d%m%Y"), ".csv"))

```


# training data 


```{r}

smart_raw_df <- read_csv(save_location)

view_df <- smart_raw_df |>
  group_by(unique_id) |>
  fill(email, .direction = "downup") %>%
  ungroup() |>
  count(email, unique_id, password, stage, str_sub(timestamp, 1, 10)) |>
  filter(str_detect(email, "vanroosbroecka@gmail.com"),
         password != "CLEARED")

smart_raw_df |>
  colnames()

```




