---
title: "processing"
author: "Jamie Cummins"
date: "3/15/2021"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(Hmisc)
library(RSQLite)
library(jsonlite)
library(tidyverse)
library(janitor)
library(pacman)
library(knitr)
library(kableExtra)
library(lubridate) 
library(RCurl)
library(threadr)
library(emayili)

```


# download data from server and save to specific location 

```{r}

## Change these:
### make sure you leave the "data.sqlite" and "data.csv" bits
download_location <- paste0('raw/training/smart_data_', format(Sys.Date(), "%d%m%Y"), '.sqlite')
save_location <- paste0('raw/training/smart_data_', format(Sys.Date(), "%d%m%Y"), '.rds')


## Change these to specify which procedure you want to download:

### Required endings of url:
#### For RAI: rai-elaborated-nl/data/data.sqlite
#### For SPM: spm-nl/data/data.sqlite
#### For self-reports: b_qol_gds/data/data.sqlite
#### For SMART: training-1/data/data.sqlite
#### For NBACK: training-2/data/data.sqlite

url <- "ftp://ftp.drjamiecumminscom.webhosting.be/www/md/training-1/data/data.sqlite"
credentials <- "mariedelabie@drjamiecumminscom:MarieDelabie1"



# download questionnaire data and save it to location
download_ftp_file(file_remote = url,
                  file_local = download_location,
                  credentials = credentials)

```

# Converting SQL data

Establish SQL connection and extract main table:

```{r}

# Establish SQL connection
con <-
  dbConnect(drv = RSQLite::SQLite(),
            # GUYS: CHANGE THE BELOW DIRECTORY TO THE LOCATION AND NAME OF THE FILE
            dbname = download_location)


# Extract main table
d <-
  dbGetQuery(conn = con,
             statement = 'SELECT * FROM labjs')

# Close and discard connection
dbDisconnect(conn = con)
rm(con)

```

* id: row number
* session: generated identifier for each individual session (one per participant)
* timestamp
* url
* metadata: also includes a variable "id" which corresponds to individual sessions
* data: the actual data

Process metadata and combine with the rest:

```{r}

d.meta <-
  map_dfr(d$metadata, fromJSON) %>% # apply JSON function to the metadata
  dplyr::rename(unique_id = id) # rename to avoid confusion

d <- d %>%
  bind_cols(d.meta) %>%
  select(-metadata)

```

* slice: the number of the entry for that specific subject
* observation: what we will use as subject identifier
* payload: tells you whether the data are complete (full) or not (incremental)

See by how much subject identifiers can be shortened:

```{r}

count_unique <- function(x) {
  return(length(unique(x)))
}
# returns how many unique observations there are in a given vector

information_preserved <- function(x, length) {
  return(count_unique(str_sub(x, end = i)) == count_unique(x))
}
# put in a vector and a string length
# for that string length, the vector checks how many unique observations there are when the strings are truncated using that length
# the function then compares this number to the number of unique observations when at full string length
# if this number is the same (i.e., truncating the string doesn't change the number of unique observations) it returns TRUE

for (i in 5:36) {
  if (
    information_preserved(d$session, i) &&
    information_preserved(d$observation, i)
  ) {
    break()
  }
}
# check if more than five characters are needed to preserve the identifying information
# for-loop breaks as soon as both expressions are evaluated as TRUE

```

Truncate ID strings to five characters:

```{r}

# Truncate IDs
d <- d %>%
  dplyr::mutate(session = str_sub(session, end = i),
                unique_id = str_sub(unique_id, end = i))

```

Extract the actual data:

```{r}

# Create function that does this
parseJSON <- function(input) {
  return(input %>%
           fromJSON(flatten = TRUE) %>% {
             if (class(.) == 'list') {discard(., is.null) %>% as_tibble()}
             else {.}
             } %>%
           janitor::clean_names() %>%
           mutate_all(as.character))
}

# Get full data
d.full <- d %>%
  dplyr::filter(payload == 'full') %>%
  group_by(unique_id, id) %>%
  do({map_dfr(.$data, parseJSON)} %>%
       bind_rows()) %>%
  ungroup()

```

Get incremental data (in case the full data are not available).

```{r}

d.incremental <-
  d %>%
  dplyr::filter(payload %in% c('incremental', 'latest')) %>%
  group_by(unique_id, id) %>%
  do(
    {map_dfr(.$data, parseJSON)} %>%
      bind_rows()
  ) %>%
  ungroup()

# merge data sets
d.output <-
  d.full %>%
  bind_rows(d.incremental %>% filter(!(unique_id %in% d.full$unique_id)))

full_raw_data_df <- d.output

```

# write out and save rds objects

```{r}

# this saves the data as a .rds file to the save_location you set at the top of the script 
write_rds(full_raw_data_df, save_location)

```


#######
Below is the code for processing the data downloaded above
#######


# process demographics

```{r}

# questionnaire_df <- questionnaire_raw_df %>%
#   filter(!is.na(wat_is_uw_e_mailadres)) %>%
#   select(email = wat_is_uw_e_mailadres,
#          age = wat_is_uw_leeftijd,
#          complaints = welke_cognitieve_klachten_heeft_u,
#          preexisting_condition = hebt_u_een_bepaalde_neurologische_of_psychiatrische_diagnose,
#          brain_injury = hebt_u_ooit_een_hersenletsel_gehad)

```


```{r}

# write_csv(questionnaire_df, paste0("processed/questionnaire_processed_data_", format(Sys.Date(), "%d%m%Y"), ".csv"))

```


# training data 


```{r}

# run the two below lines together at the same time
set.seed(42)
passwords <- data.frame(password = sample(c(1000:9999), 118),
                        stage = as.numeric(c(5:122)))

# example of a manually specified directory 
## "raw/training/smart_data_28112022.rds"
smart_raw_df <- read_rds(save_location)

view_df <- smart_raw_df |>
  group_by(unique_id) |>
  fill(email, .direction = "downup") %>%
  fill(stage, .direction = "downup") |>
  mutate(stage = as.numeric(stage)) |>
  ungroup() |>
  count(email, unique_id, stage, timestamp = str_sub(timestamp, 1, 10)) |>
  filter(str_detect(email, "emilypuck.shop@gmail.com")) |>
  select(-n) |>
  left_join(passwords)


```

# Weekly completions

The below chunk tells you how many sessions the named participants completed in the previous week. If the participant you are looking for isn't listed, then they (probably) haven't done the training in the past week. 

```{r}

smart_raw_df <- read_rds(save_location)

seven_days_ago <- as.Date(Sys.Date() - 7) 
  

weekly_df <- smart_raw_df |>
  group_by(unique_id) |>
  fill(email, .direction = "downup") |>
  fill(timestamp, .direction = "downup") |>
  filter(sender != "Welcome") |>
  ungroup() |>
  select(timestamp, email, unique_id) |>
  mutate(timestamp = as.Date(timestamp)) |>
  filter(timestamp >= seven_days_ago) |>
  distinct(email, timestamp) |>
  count(email)

```



# Actual Processing Script

```{r}

library(tidyverse)

rai_raw_df <- read_rds("raw/assessments/rai_03012023.rds") |>
  fill(prolificid, .direction = "downup") |>
  rename(email = prolificid) |>
  mutate(timestamp = lubridate::date(timestamp),
         email = tolower(email))
spm_raw_df <- read_rds("raw/assessments/spm_03012023.rds") |>
  fill(email, .direction = "downup") |>
  mutate(timestamp = lubridate::date(timestamp),
         email = tolower(email))
self_reports_raw_df <- read_rds("raw/assessments/self_reports_03012023.rds") |>
  fill(wat_is_uw_email, .direction = "downup") |>
  rename(email = wat_is_uw_email) |>
  mutate(timestamp = lubridate::date(timestamp),
         email = tolower(email))

conditions_df <- read_csv2("conditions.csv") |>
  select(email, condition)

```

# process Raven's

```{r}

complete_spm_sessions <- spm_raw_df |>
  filter(!is.na(main_image)) |>
  group_by(unique_id, timestamp) |>
  filter(n() == 60) |>
  ungroup() |>
  distinct(unique_id) |>
  as_vector()

# process Raven's Matrices
fluid_iq_df <- spm_raw_df |>
  filter(unique_id %in% complete_spm_sessions,
         !is.na(correct_answer)) |>
  mutate(correct = case_when(correct_answer == response ~ 1,
                             TRUE ~ 0)) |>
  group_by(email, timestamp) |>
  summarise(ravens_raw_score = sum(as.numeric(correct)),
            ravens_duration_minutes = sum(as.numeric(duration)) / (1000 * 60)) |>
  ungroup() |>
  mutate(fluid_iq = case_when(ravens_raw_score < 21 ~ 65,
                              ravens_raw_score > 20 & ravens_raw_score < 25 ~ 66,
                              ravens_raw_score == 25 ~ 67,
                              ravens_raw_score == 26 ~ 68,
                              ravens_raw_score == 27 ~ 69,
                              ravens_raw_score == 28 ~ 70,
                              ravens_raw_score == 29 ~ 70,
                              ravens_raw_score == 30 ~ 71,
                              ravens_raw_score == 31 ~ 74,
                              ravens_raw_score == 32 ~ 74,
                              ravens_raw_score == 33 ~ 75,
                              ravens_raw_score == 34 ~ 76,
                              ravens_raw_score == 35 ~ 77,
                              ravens_raw_score == 36 ~ 80,
                              ravens_raw_score == 37 ~ 82,
                              ravens_raw_score == 38 ~ 83,
                              ravens_raw_score == 39 ~ 85,
                              ravens_raw_score == 40 ~ 87,
                              ravens_raw_score == 41 ~ 89,
                              ravens_raw_score == 42 ~ 90,
                              ravens_raw_score == 43 ~ 91,
                              ravens_raw_score == 44 ~ 94,
                              ravens_raw_score == 45 ~ 96,
                              ravens_raw_score == 46 ~ 98,
                              ravens_raw_score == 47 ~ 100,
                              ravens_raw_score == 48 ~ 103,
                              ravens_raw_score == 49 ~ 104,
                              ravens_raw_score == 50 ~ 107,
                              ravens_raw_score == 51 ~ 109,
                              ravens_raw_score == 52 ~ 111,
                              ravens_raw_score == 53 ~ 114,
                              ravens_raw_score == 54 ~ 116,
                              ravens_raw_score == 55 ~ 120,
                              ravens_raw_score == 56 ~ 123,
                              ravens_raw_score == 57 ~ 124,
                              ravens_raw_score == 58 ~ 127,
                              ravens_raw_score == 59 ~ 130,
                              ravens_raw_score == 60 ~ 132)) %>%
  # full_join(conditions_df) %>%
  
  # need to do percentile scores properly
  mutate(ravens_percentile = round(percent_rank(as.numeric(ravens_raw_score)), 2))

# spm_complete_data <- fluid_iq_df |>
#   group_by(email) |>
#   filter(n() == 2) |>
#   ungroup() |>
#   pull(email) 

```

# process RAI

```{r}

complete_rai_sessions <- rai_raw_df |>
  filter(trial_num != "NA") |>
  group_by(unique_id, timestamp) |>
  distinct(trial_num, timestamp, .keep_all = TRUE) |>
  filter(n() == 128) |>
  ungroup() |>
  distinct(unique_id) |>
  as_vector()

relational_task_trials_df <- rai_raw_df %>%
  filter(unique_id %in% complete_rai_sessions) %>%
  group_by(email, timestamp) %>%
  fill(total_correct, .direction = "downup") %>%
  mutate(total_correct = as.numeric(lead(total_correct)),
         duration = as.numeric(duration),
         correct = as.numeric(as.logical(correct))) %>%
  ungroup() %>%
  filter(sender == "trial" & trial_num != "XXX") %>%
  select(email, timestamp, unique_id, correct, block, correct_response, response, duration, total_correct) %>%
  group_by(email, timestamp) %>%
  mutate(n_trials_completed = n(),
         rts_criterion = case_when(quantile(duration, .5) >= 5000 ~ 1,
                               TRUE ~ 0)) %>%
  mutate(trial_type = case_when(block == 1 ~ "opposition",
                                block == 2 ~ "difference",
                                block == 3 ~ "quantity",
                                block == 4 ~ "temporal",
                                block == 5 ~ "containment",
                                block == 6 ~ "analogy",
                                block == 7 ~ "deictic",
                                block == 8 | block == 9 ~ "mathematical")) %>%
  group_by(email, timestamp, trial_type) %>%
  mutate(correct = case_when(is.na(correct) ~ 0,
                             TRUE ~ correct),
         trial_score = mean(correct)) %>%
  ungroup()


# get overall score
relational_task_overall_scores_df <- relational_task_trials_df %>%
  arrange(email, timestamp, total_correct) %>%
  group_by(email, timestamp) %>%
  summarise_all(last) %>%
  mutate(relational_score = (total_correct / n_trials_completed)) %>%
  ungroup() %>%
  select(email, timestamp, relational_score, rts_criterion)

# rai_complete_data <- relational_task_overall_scores_df %>%
#   group_by(email, timepoint) %>%
#   filter(n() == 2) %>%
#   ungroup() %>%
#   pull(email) 

```


# combined df

```{r}

group_level_df <- fluid_iq_df |>
  full_join(relational_task_overall_scores_df) |>
  full_join(conditions_df) |>
  group_by(email) |>
  filter(n() == 2)

```
