---
title: "processing"
author: "Jamie Cummins"
date: "3/15/2021"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(Hmisc)
library(RSQLite)
library(jsonlite)
library(tidyverse)
library(janitor)
library(pacman)
library(knitr)
library(kableExtra)
library(lubridate) 
library(RCurl)
library(threadr)
library(emayili)

```


# download data from server and save to specific location 

```{r}

## Change these:
### make sure you leave the "data.sqlite" and "data.csv" bits
download_location <- paste0('raw/assessments/rai_', format(Sys.Date(), "%d%m%Y"), '.sqlite')
save_location <- paste0('raw/assessments/rai_', format(Sys.Date(), "%d%m%Y"), '.rds')


## Change these to specify which procedure you want to download:

### Required endings of url:
#### For RAI: rai-elaborated-nl/data/data.sqlite
#### For SPM: spm-nl/data/data.sqlite
#### For self-reports: b_qol_gds/data/data.sqlite
#### For SMART: training-1/data/data.sqlite
#### For NBACK: training-2/data/data.sqlite


url <- "ftp://ftp.drjamiecumminscom.webhosting.be/www/md/rai-elaborated-nl/data/data.sqlite"
credentials <- "mariedelabie@drjamiecumminscom:MarieDelabie1"



# download questionnaire data and save it to location
download_ftp_file(file_remote = url,
                  file_local = download_location,
                  credentials = credentials)

```

# Converting SQL data

Establish SQL connection and extract main table:

```{r}

# Establish SQL connection
con <-
  dbConnect(drv = RSQLite::SQLite(),
            # GUYS: CHANGE THE BELOW DIRECTORY TO THE LOCATION AND NAME OF THE FILE
            dbname = download_location)


# Extract main table
d <-
  dbGetQuery(conn = con,
             statement = 'SELECT * FROM labjs')

# Close and discard connection
dbDisconnect(conn = con)
rm(con)

```

* id: row number
* session: generated identifier for each individual session (one per participant)
* timestamp
* url
* metadata: also includes a variable "id" which corresponds to individual sessions
* data: the actual data

Process metadata and combine with the rest:

```{r}

d.meta <-
  map_dfr(d$metadata, fromJSON) %>% # apply JSON function to the metadata
  dplyr::rename(unique_id = id) # rename to avoid confusion

d <- d %>%
  bind_cols(d.meta) %>%
  select(-metadata)

```

* slice: the number of the entry for that specific subject
* observation: what we will use as subject identifier
* payload: tells you whether the data are complete (full) or not (incremental)

See by how much subject identifiers can be shortened:

```{r}

count_unique <- function(x) {
  return(length(unique(x)))
}
# returns how many unique observations there are in a given vector

information_preserved <- function(x, length) {
  return(count_unique(str_sub(x, end = i)) == count_unique(x))
}
# put in a vector and a string length
# for that string length, the vector checks how many unique observations there are when the strings are truncated using that length
# the function then compares this number to the number of unique observations when at full string length
# if this number is the same (i.e., truncating the string doesn't change the number of unique observations) it returns TRUE

for (i in 5:36) {
  if (
    information_preserved(d$session, i) &&
    information_preserved(d$observation, i)
  ) {
    break()
  }
}
# check if more than five characters are needed to preserve the identifying information
# for-loop breaks as soon as both expressions are evaluated as TRUE

```

Truncate ID strings to five characters:

```{r}

# Truncate IDs
d <- d %>%
  dplyr::mutate(session = str_sub(session, end = i),
                unique_id = str_sub(unique_id, end = i))

```

Extract the actual data:

```{r}

# Create function that does this
parseJSON <- function(input) {
  return(input %>%
           fromJSON(flatten = TRUE) %>% {
             if (class(.) == 'list') {discard(., is.null) %>% as_tibble()}
             else {.}
             } %>%
           janitor::clean_names() %>%
           mutate_all(as.character))
}

# Get full data
d.full <- d %>%
  dplyr::filter(payload == 'full') %>%
  group_by(unique_id, id) %>%
  do({map_dfr(.$data, parseJSON)} %>%
       bind_rows()) %>%
  ungroup()

```

Get incremental data (in case the full data are not available).

```{r}

d.incremental <-
  d %>%
  dplyr::filter(payload %in% c('incremental', 'latest')) %>%
  group_by(unique_id, id) %>%
  do(
    {map_dfr(.$data, parseJSON)} %>%
      bind_rows()
  ) %>%
  ungroup()

# merge data sets
d.output <-
  d.full %>%
  bind_rows(d.incremental %>% filter(!(unique_id %in% d.full$unique_id)))

full_raw_data_df <- d.output

```

# write out and save csvs

```{r}

# CHANGE LOCATION AND NAME IN THE STRING TO BE WHERE YOU WANT TO SAVE THE FILE
write_rds(full_raw_data_df, save_location)

```


Notes to self:

1. Go through SMART data -> identify participants who completed only partial training but also gave follow-up measures, then run sensitivity analyses
2. Extract descriptive data about the number of sessions that each participant completed
3. flag and !!probably!! include participants who only completed T1 assessments -> sensitivity analysis
4. Check for bugs in the exclusion of participants at T2
5. Check out the potential for random responding in RAI/other measures


# process demographics

```{r}

# questionnaire_df <- questionnaire_raw_df %>%
#   filter(!is.na(wat_is_uw_e_mailadres)) %>%
#   select(email = wat_is_uw_e_mailadres,
#          age = wat_is_uw_leeftijd,
#          complaints = welke_cognitieve_klachten_heeft_u,
#          preexisting_condition = hebt_u_een_bepaalde_neurologische_of_psychiatrische_diagnose,
#          brain_injury = hebt_u_ooit_een_hersenletsel_gehad)

```


```{r}

# write_csv(questionnaire_df, paste0("processed/questionnaire_processed_data_", format(Sys.Date(), "%d%m%Y"), ".csv"))

```


# training data 


```{r}

smart_raw_df <- read_rds(save_location)

view_df <- smart_raw_df |>
  group_by(unique_id) |>
  fill(email, .direction = "downup") %>%
  ungroup() |>
  count(email, unique_id, password, stage, str_sub(timestamp, 1, 10)) |>
  filter(str_detect(email, "emilypuck.shop@gmail.com"),
         password != "CLEARED")

smart_raw_df |>
  colnames()

```



# Actual Processing Script

```{r}

library(tidyverse)

rai_raw_df <- read_rds("raw/assessments/rai_03012023.rds") |>
  fill(prolificid, .direction = "downup") |>
  rename(email = prolificid) |>
  mutate(timestamp = lubridate::date(timestamp),
         email = tolower(email))
spm_raw_df <- read_rds("raw/assessments/spm_03012023.rds") |>
  fill(email, .direction = "downup") |>
  mutate(timestamp = lubridate::date(timestamp),
         email = tolower(email))
self_reports_raw_df <- read_rds("raw/assessments/self_reports_03012023.rds") |>
  fill(wat_is_uw_email, .direction = "downup") |>
  rename(email = wat_is_uw_email) |>
  mutate(timestamp = lubridate::date(timestamp),
         email = tolower(email))

conditions_df <- read_csv2("conditions.csv") |>
  select(email, condition)

```

# process Raven's

```{r}

complete_spm_sessions <- spm_raw_df |>
  filter(!is.na(main_image)) |>
  group_by(unique_id, timestamp) |>
  filter(n() == 60) |>
  ungroup() |>
  distinct(unique_id) |>
  as_vector()

# process Raven's Matrices
fluid_iq_df <- spm_raw_df |>
  filter(unique_id %in% complete_spm_sessions,
         !is.na(correct_answer)) |>
  mutate(correct = case_when(correct_answer == response ~ 1,
                             TRUE ~ 0)) |>
  group_by(email, timestamp) |>
  summarise(ravens_raw_score = sum(as.numeric(correct)),
            ravens_duration_minutes = sum(as.numeric(duration)) / (1000 * 60)) |>
  ungroup() |>
  mutate(fluid_iq = case_when(ravens_raw_score < 21 ~ 65,
                              ravens_raw_score > 20 & ravens_raw_score < 25 ~ 66,
                              ravens_raw_score == 25 ~ 67,
                              ravens_raw_score == 26 ~ 68,
                              ravens_raw_score == 27 ~ 69,
                              ravens_raw_score == 28 ~ 70,
                              ravens_raw_score == 29 ~ 70,
                              ravens_raw_score == 30 ~ 71,
                              ravens_raw_score == 31 ~ 74,
                              ravens_raw_score == 32 ~ 74,
                              ravens_raw_score == 33 ~ 75,
                              ravens_raw_score == 34 ~ 76,
                              ravens_raw_score == 35 ~ 77,
                              ravens_raw_score == 36 ~ 80,
                              ravens_raw_score == 37 ~ 82,
                              ravens_raw_score == 38 ~ 83,
                              ravens_raw_score == 39 ~ 85,
                              ravens_raw_score == 40 ~ 87,
                              ravens_raw_score == 41 ~ 89,
                              ravens_raw_score == 42 ~ 90,
                              ravens_raw_score == 43 ~ 91,
                              ravens_raw_score == 44 ~ 94,
                              ravens_raw_score == 45 ~ 96,
                              ravens_raw_score == 46 ~ 98,
                              ravens_raw_score == 47 ~ 100,
                              ravens_raw_score == 48 ~ 103,
                              ravens_raw_score == 49 ~ 104,
                              ravens_raw_score == 50 ~ 107,
                              ravens_raw_score == 51 ~ 109,
                              ravens_raw_score == 52 ~ 111,
                              ravens_raw_score == 53 ~ 114,
                              ravens_raw_score == 54 ~ 116,
                              ravens_raw_score == 55 ~ 120,
                              ravens_raw_score == 56 ~ 123,
                              ravens_raw_score == 57 ~ 124,
                              ravens_raw_score == 58 ~ 127,
                              ravens_raw_score == 59 ~ 130,
                              ravens_raw_score == 60 ~ 132)) %>%
  # full_join(conditions_df) %>%
  
  # need to do percentile scores properly
  mutate(ravens_percentile = round(percent_rank(as.numeric(ravens_raw_score)), 2))

# spm_complete_data <- fluid_iq_df |>
#   group_by(email) |>
#   filter(n() == 2) |>
#   ungroup() |>
#   pull(email) 

```

# process RAI

```{r}

complete_rai_sessions <- rai_raw_df |>
  filter(trial_num != "NA") |>
  group_by(unique_id, timestamp) |>
  distinct(trial_num, timestamp, .keep_all = TRUE) |>
  filter(n() == 128) |>
  ungroup() |>
  distinct(unique_id) |>
  as_vector()

relational_task_trials_df <- rai_raw_df %>%
  filter(unique_id %in% complete_rai_sessions) %>%
  group_by(email, timestamp) %>%
  fill(total_correct, .direction = "downup") %>%
  mutate(total_correct = as.numeric(lead(total_correct)),
         duration = as.numeric(duration),
         correct = as.numeric(as.logical(correct))) %>%
  ungroup() %>%
  filter(sender == "trial" & trial_num != "XXX") %>%
  select(email, timestamp, unique_id, correct, block, correct_response, response, duration, total_correct) %>%
  group_by(email, timestamp) %>%
  mutate(n_trials_completed = n(),
         rts_criterion = case_when(quantile(duration, .5) >= 5000 ~ 1,
                               TRUE ~ 0)) %>%
  mutate(trial_type = case_when(block == 1 ~ "opposition",
                                block == 2 ~ "difference",
                                block == 3 ~ "quantity",
                                block == 4 ~ "temporal",
                                block == 5 ~ "containment",
                                block == 6 ~ "analogy",
                                block == 7 ~ "deictic",
                                block == 8 | block == 9 ~ "mathematical")) %>%
  group_by(email, timestamp, trial_type) %>%
  mutate(correct = case_when(is.na(correct) ~ 0,
                             TRUE ~ correct),
         trial_score = mean(correct)) %>%
  ungroup()


# get overall score
relational_task_overall_scores_df <- relational_task_trials_df %>%
  arrange(email, timestamp, total_correct) %>%
  group_by(email, timestamp) %>%
  summarise_all(last) %>%
  mutate(relational_score = (total_correct / n_trials_completed)) %>%
  ungroup() %>%
  select(email, timestamp, relational_score, rts_criterion)

rai_complete_data <- relational_task_overall_scores_df %>%
  group_by(email) %>%
  filter(n() == 2) %>%
  ungroup() %>%
  pull(email) 

```


# combined df

```{r}

group_level_df <- fluid_iq_df |>
  full_join(relational_task_overall_scores_df) |>
  full_join(conditions_df) |>
  group_by(email) |>
  filter(n() == 2)



```
