---
title: "processing_assessments"
author: "Jamie Cummins"
date: "`r Sys.Date()`"
output: html_document
---
---
title: "processing"
author: "Jamie Cummins"
date: "3/15/2021"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(Hmisc)
library(RSQLite)
library(jsonlite)
library(tidyverse)
library(janitor)
library(pacman)
library(knitr)
library(kableExtra)
library(lubridate) 
library(tidylog)

```


```{r}

# spm t1 
spm_t1_raw_df <- read_csv("raw/spm_data_06062022.csv")

# rai t1 
rai_t1_raw_df <- read_csv("raw/rai_data_06062022.csv")

# questionnaires t1 
questionnaires_t1_raw_df <- read_csv("raw/questionnaires_data_06062022.csv")


```

# Demographics

```{r}

demographics_df <- rai_t1_raw_df %>%
  filter(!is.na(wat_is_uw_geslacht),
         age > 50,
         prolificid != "debackerinne@hotmail.com") %>%
  select(email = prolificid, gender = wat_is_uw_geslacht, age)

```

# Raven's

```{r}

spm_df <- spm_t1_raw_df %>%
  select(email, correct_answer, response) %>%
  fill(email) %>%
  filter(!is.na(correct_answer),
         email != "debackerinne@hotmail.com") %>%
  mutate(correct = case_when(correct_answer == response ~ 1,
                             TRUE ~ 0)) %>%
  select(-correct_answer, -response) %>%
  group_by(email) %>%
  filter(n() == 60) %>%
  summarise(ravens_raw_score = sum(correct)) 

```

# RAI

```{r}

rai_t1_raw_df %>%
  # select(prolificid, trial_num, correct) %>%
  fill(prolificid) %>%
  filter(prolificid == "globetrotster@hotmail.com")

rai_df <- rai_t1_raw_df %>%
  select(email = prolificid, trial_num, correct) %>%
  fill(email) %>%
  group_by(email, trial_num) %>%
  filter(n() == 4) %>%
  fill(correct) %>%
  ungroup() %>%
  mutate(correct = case_when(is.na(correct) ~ FALSE,
                             TRUE ~ correct)) %>%
  distinct(email, trial_num, .keep_all = TRUE) %>%
  select(email, correct) %>%
  group_by(email) %>%
  filter(n() == 128) %>%
  summarise(rai_score = sum(correct))

```

# Questionnaires

```{r}

# BRIEF
brief_df <- questionnaires_t1_raw_df %>%
  select(email = wat_is_uw_email, contains("brief")) %>%
  fill(email) %>%
  filter(!is.na(brief_brief_sr_1))

# EURO QOL
qol_df <- questionnaires_t1_raw_df %>%
  select(email = wat_is_uw_email, mobility = mob_mobiliteit,
         self_care = zelfzorg_zelfzorg, daily_function = dagelijkse_dagelijkse,
         pain = pijn_pijn, fear = angst_angst) %>% 
  fill(email) %>%
  pivot_longer(cols = mobility:fear,
               names_to = "item",
               values_to = "rating") %>%
  filter(!is.na(rating)) %>%
  group_by(email) %>%
  mutate(qol_score = sum(rating)) %>%
  distinct(email, qol_score)

# GDS
gds_df <- questionnaires_t1_raw_df %>%
  select(email = wat_is_uw_email, contains("gds")) %>% 
  fill(email) %>%
  filter(!is.na(gds_1_r)) %>%
  
  # reverse scoring
  mutate_at(vars(contains("_r")), 
            funs(case_when(. == 1 ~ 0,
                 TRUE ~ 1))) %>%
  group_by(email) %>%
  mutate(gds_score = sum(across(contains("gds")))) %>%
  ungroup() %>%
  select(email, gds_score)

```


# Add RBANS here

```{r}




```



```{r}

group_level_df <- spm_df %>%
  left_join(rai_df) %>%
  left_join(gds_df) %>%
  left_join(brief_df) %>%
  left_join(qol_df) 

```

# Converting SQL data

Establish SQL connection and extract main table:

```{r}

# Establish SQL connection
con <-
  dbConnect(drv = RSQLite::SQLite(),
            dbname = paste0('raw/spm_data_', format(Sys.Date(), "%d%m%Y"), '.sqlite'))


# Extract main table
d <-
  dbGetQuery(conn = con,
             statement = 'SELECT * FROM labjs')

# Close and discard connection
dbDisconnect(conn = con)
rm(con)

```

* id: row number
* session: generated identifier for each individual session (one per participant)
* timestamp
* url
* metadata: also includes a variable "id" which corresponds to individual sessions
* data: the actual data

Process metadata and combine with the rest:

```{r}

d.meta <-
  map_dfr(d$metadata, fromJSON) %>% # apply JSON function to the metadata
  dplyr::rename(unique_id = id) # rename to avoid confusion

d <- d %>%
  bind_cols(d.meta) %>%
  select(-metadata)

```

* slice: the number of the entry for that specific subject
* observation: what we will use as subject identifier
* payload: tells you whether the data are complete (full) or not (incremental)

See by how much subject identifiers can be shortened:

```{r}

count_unique <- function(x) {
  return(length(unique(x)))
}
# returns how many unique observations there are in a given vector

information_preserved <- function(x, length) {
  return(count_unique(str_sub(x, end = i)) == count_unique(x))
}
# put in a vector and a string length
# for that string length, the vector checks how many unique observations there are when the strings are truncated using that length
# the function then compares this number to the number of unique observations when at full string length
# if this number is the same (i.e., truncating the string doesn't change the number of unique observations) it returns TRUE

for (i in 5:36) {
  if (
    information_preserved(d$session, i) &&
    information_preserved(d$observation, i)
  ) {
    break()
  }
}
# check if more than five characters are needed to preserve the identifying information
# for-loop breaks as soon as both expressions are evaluated as TRUE

```

Truncate ID strings to five characters:

```{r}

# Truncate IDs
d <- d %>%
  dplyr::mutate(session = str_sub(session, end = i),
                unique_id = str_sub(unique_id, end = i))

```

Extract the actual data:

```{r}

# Create function that does this
parseJSON <- function(input) {
  return(input %>%
           fromJSON(flatten = TRUE) %>% {
             if (class(.) == 'list') {discard(., is.null) %>% as_tibble()}
             else {.}
             } %>%
           janitor::clean_names() %>%
           mutate_all(as.character))
}

# Get full data
d.full <- d %>%
  dplyr::filter(payload == 'full') %>%
  group_by(unique_id, id) %>%
  do({map_dfr(.$data, parseJSON)} %>%
       bind_rows()) %>%
  ungroup()

```

Get incremental data (in case the full data are not available).

```{r}

d.incremental <-
  d %>%
  dplyr::filter(payload %in% c('incremental', 'latest')) %>%
  group_by(unique_id, id) %>%
  do(
    {map_dfr(.$data, parseJSON)} %>%
      bind_rows()
  ) %>%
  ungroup()

# # merge data sets
# d.output <-
#   d.full %>%
#   bind_rows(d.incremental %>% filter(!(unique_id %in% d.full$unique_id)))

# %>%
#   type_convert()

full_raw_data_df <- d.incremental

```

# write out

```{r}

write_csv(full_raw_data_df, paste0('raw/spm_data_', format(Sys.Date(), "%d%m%Y"), '.csv'))

```